An Experimental Evaluation of Hybrid Querying on Vectors (EA\&B)

### 2025-07-22 Revision/response.tex

%%%%%%%%%%%%%%%%%%%%%% R1 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{metarevision}
		\textit{Improve the readability of some figures (e.g., performance under different attribute distributions or selectivity).}
	\end{metarevision}
}

\noindent
\textbf{Response:} We sincerely appreciate your valuable suggestions, which provide meaningful improvements to our work. Ensuring the readability of the figures is indeed crucial. We modified both the figure and the related description in the paper to improve readability and ensure academic rigor. We have revised the manuscript accordingly. (Please see Section 4.2.3, highlighted in blue in the revised manuscript)

For performance under different attribute distributions, to illustrate how different algorithms are affected by varying attribute distributions, we have added robustness indicators (Strong and Weak) in the figure. This provides more intuitive visualization of how robust each algorithm performs across different attribute distributions. As shown in Figure\ref{fig:attribute_distribution} below.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{fig/attribute_distribution.png}
	\caption{}
	\label{fig:attribute_distribution}
\end{figure}

@@

For performance under different attribute selectivity, to reflect the performance differences of each algorithm under varying attribute selectivities, we have added performance rankings for each algorithm across different selectivity levels in the figure. This provides a clearer illustration of how each algorithm is affected by attribute selectivity rates. As shown in Figure\ref{fig:attribute_selectivity} below.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{fig/attribute_selectivity.png}
	\caption{}
	\label{fig:attribute_selectivity}
\end{figure}

@@

\textcolor{blue}{
		"Similar to attribute filtering, the performance of multimodal queries with range filtering shown in Figure 16a is also consistent with the single-modality results in Figure 14. As shown in Figure 15, the algorithms still perform the worst on multimodal datasets like Text2Image. This is because these range query algorithms are almost graph-based. Graph algorithms need more iterations to find true neighbors in OOD datasets."
}

%%%%%%%%%%%%%%%%%%%%%% R4 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{metarevision}
		\textit{Instead of using the point ID, run experiments with meaningful attributes for RF-NN search on datasets (Deep, YT-Audio, WIT).}
	\end{metarevision}
}

\noindent
\textbf{Response:} 
Thank you for your suggestion. Using IDs as attributes for range filtering is a prevalent method in current range filtering algorithm evaluations, which indeed has considerable limitations. To address this issue, we have added two datasets that use real attributes:

\textbf{YT-Audio-Real dataset:} For this dataset, we utilized the URLs provided by the YouTube8M dataset to crawl corresponding video categories and view counts as real attributes, using \textbf{video categories as attributes for attribute filtering experiments} and \textbf{view counts as attributes for range filtering experiments}, respectively.

\textbf{WIT-Real dataset:} We used the \texttt{clip-vit-base-patch32} model to encode the \texttt{page\_title} field from the WIT dataset into vectors, and extracted the language type and the length of the \texttt{context\_page\_description} field as real attributes, applying them in both attribute filtering and range filtering experiments.

@@

\noindent
\textbf{Response:} 
That's a great suggestion. In order to figure out if the server setting affects the relative ordering of methods, we conducted additional experiments on a server equipped with two Intel(R) Xeon(R) Gold 6330 CPUs @ 2.00GHz using the YT-Audio-Real dataset.
The results show that the performance differences across algorithms remain small, and their relative rankings do not change significantly across servers.
We did observe slightly lower QPS on this server compared to our original one, which is mainly due to the lower CPU frequency of the Xeon 6330 processors. Please see section 4.2.4.

\textcolor{blue}{"To investigate the impact of different server configurations on the performance of the algorithms, we conducted additional experiments on a server equipped with two Intel(R) Xeon(R) Gold 6330 CPUs @ 2.00GHz using the YT-Audio-Real dataset, and the results are shown in Figure 12(b). Compared with the experimental results on the original server (as shown in Figure 4d), the relative ranking of the algorithms remains largely unchanged, indicating that the performance of these methods is not significantly affected by differences in CPU configurations. Due to the lower CPU frequency of the Xeon 6330 compared to the original server, the QPS values are slightly lower across all methods."}

%%%%%%%%%%%%%%%%%%%%%% R6 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{metarevision}
		\textit{In Table 1, please clarify which methods are disk-based, and which methods are memory-based.
			For disk-based methods, it is also meaningful to report the cost breakdown (e.g., in terms of I/O time and CPU time).}
	\end{metarevision}
}

\noindent
\textbf{Response:} 
Thank you for your suggestion! Currently, except for FilteredDiskANN, other hybrid query methods are memory-based, so we added the feature of whether to support disk for each algorithm in Table 1. At the same time, we also compared the search performance difference of FilteredDiskANN on memory and SSD using real data sets. The results are shown in Figure\ref{fig:recall-latency}:
\begin{figure}[htbp]
	\centering
	\includegraphics[width=\linewidth]{fig/recall_latency_all.png}
	\caption{}
	\label{fig:recall-latency}
\end{figure}

Due to space limitations and the fact that only one method supports disk-based execution, we provide the breakdown of I/O and CPU costs for disk-based methods in the following analysis. To this end, we conducted additional experiments on the real-world YT8M dataset, comparing three variants of DiskANN.

\begin{itemize}
	\item SSD-based FilteredDiskANN (original disk-based method)
	\item FilteredVamana (memory-based method)
	\item StitchedVamana (memory-based method)
\end{itemize}

@@

All experiments were run with 32 search threads for fair comparison. The following figure shows the mean latency (log-scale, left y-axis) and I/O vs CPU proportion (right y-axis) across different Recall\@10 levels:

\textbf{ Key Observations:}

Latency-Recall Tradeoff:
As expected, SSD-based FilteredDiskANN exhibits significantly higher latency than in-memory methods, especially at higher recall levels. When Recall@10 reaches 0.96, its latency increases super-linearly to over 100 ms.

@@

Cost Breakdown of DiskANN:The I/O proportion consistently dominates, maintaining 80\%–85\% of the total latency even at different recall levels.
The CPU proportion remains low (10–15\%), indicating that I/O is the major bottleneck in disk-based search pipelines.
This clearly shows that improving disk read efficiency or reducing I/O volume is critical for performance optimization.

\textbf{In-memory Methods:}

FilteredVamana achieves a good balance, staying under 30 ms while achieving Recall@10 > 0.97.
While StitchedVamana achieves extremely low latency (sub-5 ms) due to its stitched multi-label index structure, its fixed edge connectivity may limit effectiveness under highly uncorrelated label distributions. Nevertheless, it consistently supports 90\%+ Recall@10 across diverse datasets, as shown in our evaluations.

@@

%%%%%%%%%%%%%%%%%%%%%% R7 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{metarevision}
		\textit{Report the peak memory usage of query processing.}
	\end{metarevision}
}

\noindent
\textbf{Response:} 
Thank you for your suggestion. We are deeply aware of this deficiency, so we added the search peak memory usage of each algorithm in \textcolor{blue}{Section 4.2.1} of attribute filtering and \textcolor{blue}{Section 4.3.1}  of range filtering, and carefully analyzed the reasons.

\textit{\textbf{ \textcolor{blue}{" Search peak memory.}}}

\textcolor{blue}{As shown in Figure, we can find that the peak memory usage of each algorithm during search does not differ significantly. Among them, CAPS and NHQ have relatively higher peak memory usage, while the database's peak memory usage is comparatively lower."}

\textit{\textbf{ \textcolor{blue}{"Search peak memory.}}}

\textcolor{blue}{As for the Search peak memory usage, we can find from the Figure that it is basically consistent with the Build peak memory usage trend. The algorithm needs to load the complete index into memory during search. Therefore, algorithms with high building overhead, such as WinFilter, also have high search overhead; algorithms with low building overhead, such as Faiss, also have low search overhead. For Milvus, search does not require full index loading and there is additional metadata overhead when building, so the performance is different from the trend."}

@@

\noindent
\textbf{Response:}

\setcounter{detailed}{0}

%%%%%%%%%%%%%%%%%%%%%% D1 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{detailed}
		\textit{Section 2.2. Graph based and IVF are well known. Missing in the index description is the implementation of attributes for AF and RF as the section is call hybrid ANN.}
	\end{detailed}
}

\noindent
\textbf{Response:}
Thank you for your suggestion. We acknowledge that the original manuscript did not clearly explain how attribute filtering (AF) and range filtering (RF) are embedded into standard graph-based and IVF-based index structures. We have revised Section 2.2 and added two paragraphs. These paragraphs summarize how the algorithms discussed in this paper implement hybrid query functionality based on these foundational indexing structures.

\textcolor{blue}{"Graph-based indexes are widely adopted in Hybrid ANN search algorithms due to their excellent performance. To support hybrid queries, existing approaches usually augment the original graph structures by explicitly associating attribute or range information with vertices or edges. In attribute filtering, a common approach is to attach attribute metadata to vertices or edges, allowing the traversal process to skip nodes or edges that do not meet the query's attribute conditions (e.g., NHQ, Filtered-DiskANN, ACORN). In range filtering, edges typically store valid attribute intervals, and traversal proceeds only along edges satisfying the range constraints, effectively narrowing down the search space (e.g., SeRF, DSG)."}

\textcolor{blue}{"IVF-based indexes also offer unique advantages in Hybrid ANN search scenarios, such as naturally supporting pre-filtering and requiring significantly lower memory consumption and smaller index sizes compared to graph-based indexes. With minor modifications, IVF indexes achieve attribute and range filtering by directly excluding vectors that do not meet query constraints through pre-filtering, thus skipping irrelevant vectors during intra-cluster distance computations (e.g., Faiss, Milvus). Moreover, clusters can be further partitioned hierarchically based on attributes (hierarchical partitioning), resulting in finer-grained sub-partitions that significantly enhance the efficiency of attribute filtering (e.g., CAPS, PUCK)."}

### 2025-07-22 Revision/response.tex

\noindent
\textbf{Response:}

\section*{Response to Review 1}

%%%%%%%%%%%%%%%%%%%%%% W1 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{weakness}
		\textit{Although the authors justify the synthetic attribute generation, real-world applications may involve more complex attribute semantics and correlations, which might affect generalizability.}
	\end{weakness}
}

\noindent
\textbf{Response:} I really appreciate your suggestion. In real-world applications, vectors and their attributes often have complex relationships. To better reflect such scenarios, we have introduced two real-world datasets, which are described in detail in Meta-Review R3.
Interestingly, our results indicate that using real attribute values has limited impact on the overall performance trends.(Please see the experimental results in Section 4.2.2 and Section 4.3.2)

@@

%%%%%%%%%%%%%%%%%%%%%% W4 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{weakness}
		\textit{The authors have reported the peak memory usage of index construction only.
			They are also recommended to report the peak memory usage of query processing.}
	\end{weakness}
}

\noindent
\textbf{Response:} 
Thank you for your suggestion! We have made a detailed response to this question in R6 of Response To Meta-reviewer. Please see R6 for details.

%%%%%%%%%%%%%%%%%%%%%% D1 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{detailed}
		\textit{The authors have cited an experimental paper [49],
			but have not discussed about the main difference between their paper and [49].}
	\end{detailed}
}

\noindent
\textbf{Response:} 
Thank you for your suggestion! The paper [49] is a comprehensive survey and evaluation of graph structure approximate nearest neighbor search algorithms. Our paper extends this to a more complex "hybrid query" scenario and evaluates approximate search algorithms that combine vector similarity and attribute filtering/range filtering. Since hybrid queries are developed based on ANNs, we refer to the metrics and evaluation architectures of different methods proposed in this paper.

### 2025-07-22 Revision/response.tex

\noindent
\textbf{Response:} 
Thank you for your valuable suggestion regarding our experiments. In Section 4.2.4, we add experiments on the Deep100M dataset. The results show that most algorithms struggle to balance performance with time and space costs on this large-scale dataset. In particular, StitchedVamana performs poorly on Deep100M. Please see section 4.2.4.

\textcolor{blue}{"On the 100M dataset, UNG and NHQ still exhibit excellent performance, demonstrating their good adaptability to large-scale datasets; however, this also comes with significant memory overhead. The clustering-based algorithm \texttt{Faiss+HQI\_Batch} shows moderate performance but has very low memory overhead, which makes it highly attractive for practical applications. We did not present the experimental results for StitchedVamana because we found that its recall consistently remained at a very low level. Its search starting point is set to 0 by default, and when the entry point is far from the target area, it easily gets stuck in a local optimum, making it difficult to navigate to the vicinity of the correct point."}

%%%%%%%%%%%%%%%%%%%%%% R9 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{metarevision}
		\textit{Clearer insights and guidelines wrt the strength and weaknesses of the compared system.}
	\end{metarevision}
}

\noindent
\textbf{Response:} 
Thank you for your valuable suggestion. We agree that our previous analysis of the strengths and weaknesses of the algorithms was not sufficiently clear.

To address this, we add a summary table that provides a clearer and more structured comparison of the strengths and weaknesses of each algorithm, offering deeper insights and more concrete guidelines for understanding their trade-offs.
%下面放那个优劣势表

@@

\noindent
\textbf{Response:}

\section*{Response to Review 1}

%%%%%%%%%%%%%%%%%%%%%% W1 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{weakness}
		\textit{Although the authors justify the synthetic attribute generation, real-world applications may involve more complex attribute semantics and correlations, which might affect generalizability.}
	\end{weakness}
}

\noindent
\textbf{Response:} Thank you for your suggestion! In real-world applications, there may indeed exist more complex semantic correlations between vectors and their associated attributes or among different attributes themselves. To better reflect such scenarios, we introduced two real-world datasets, where both vectors and attributes are derived from actual data. For example, in the YouTube dataset, the attributes include video categories and view counts, while in the WIT dataset, the attributes correspond to the language of image descriptions.

We conducted experiments on both datasets and present the performance analysis for various algorithms in \textcolor{blue}{Sections 4.2.2 and 4.3.2}. Interestingly, our results indicate that using real attribute values has limited impact on the overall performance trends.

@@

%%%%%%%%%%%%%%%%%%%%%% D2 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{detailed}
		\textit{In Section 1.1, I could not find convincing references for the requirements in practical applications.
			[12] and [36] are research papers only, but not from real applications.}
	\end{detailed}
}

\noindent
\textbf{Response:} 
Thank you for your suggestion. We cited these two papers in the introduction to show that there are several papers that study attribute distribution and selectivity.
In addition, attribute distribution and selectivity are indeed widely used in real-world scenarios. For attribute distribution, the label attributes of many recommendation application scenarios in reality conform to the power-law distribution. For attribute selectivity: For example, if you need to query data with a very small proportion of a certain attribute, choosing a better algorithm will greatly improve the search effect. There are also some databases with practical applications that mention attribute selectivity, such as Milvus and AnalyticDB-V. For the two parts you suggested, we will add more comprehensive references in the paper.
Please see section 1.1. \textcolor{blue}{"For attribute filtering algorithms, factors such as the number of attributes, attribute distribution[12], and attribute  selectivity [46,51] significantly impact algorithm performance."}

%%%%%%%%%%%%%%%%%%%%%% D3 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{detailed}
		\textit{In Definition 2.3, $ s_1 $ and $ s_n $ should be replaced with $ a_1 $ and $ a_n $ , respectively.}
	\end{detailed}
}

\noindent
\textbf{Response:} 
Thank you for pointing it out. We sincerely apologize for the misuse of notation in Definition 2.3. This has been corrected in the revised version. (Please see Definition 2.3 in Section 2.1, highlighted in blue in the revised manuscript)

@@

\textcolor{blue}{Let \( D = \{(v_1, a_1), \ldots, (v_n, a_n)\} \) be a dataset of \( n \) \( d \)-dimensional vectors, each associated with an attribute value \( a_i \).}

%%%%%%%%%%%%%%%%%%%%%% D4 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{detailed}
		\textit{ In Section 4.1, please clarify the URL of the implementation of each method.}
	\end{detailed}
}

\noindent
\textbf{Response:} 
Thank you for the suggestion. Since there are multiple methods, listing all URLs in Section 4.1 would be overwhelming. Instead, we add the corresponding URLs in footnotes on the relevant pages throughout the paper. (Please see the relevant footnotes in Section 3 in the revised manuscript)

### 2025-07-23 Revision/response.tex

\noindent
\textbf{Response:} 
Thank you for pointing it out. We sincerely apologize for the misuse of notation in Definition 2.3. This has been corrected in the revised version. (Please see Definition 2.3 in Section 2.1, highlighted in blue in the revised manuscript)

\textcolor{blue}{Let \( D = \{(v_1, a_1), \ldots, (v_n, a_n)\} \) be a dataset of \( n \) \( d \)-dimensional vectors, each associated with an attribute value \( a_i \).}

%%%%%%%%%%%%%%%%%%%%%% D4 %%%%%%%%%%%%%%%%%%%%%%
\myfbox{
	\begin{detailed}
		\textit{ In Section 4.1, please clarify the URL of the implementation of each method.}
	\end{detailed}
}

\noindent
\textbf{Response:} 
Thank you for the suggestion. Since there are multiple methods, listing all URLs in Section 4.1 would be overwhelming. Instead, in Section 3, we add superscript markers next to each algorithm name and provide the corresponding implementation URLs as footnotes on the same page. (Please see the relevant footnotes in Section 3 in the revised manuscript)

### 2025-07-23 VLDB2026/main.tex

		%\vspace{0.1cm}
		\begin{subfigure}[t]{0.507\columnwidth}
			\centering
			\captionsetup{font=small}
			\includegraphics[width=\linewidth]{figures/exp/attribute_multimodel.pdf}
			\caption{\footnotesize MultiModel}
			\label{fig:MultiModel}
		\end{subfigure}	
		\hfill
		\begin{subfigure}[t]{0.48\columnwidth}
			\centering
			\captionsetup{font=small}
			\includegraphics[width=\linewidth]{figures/exp/attribute_multimodel_1.pdf}
			\caption{\footnotesize MultiModel}
			\label{fig:MultiModel}
		\end{subfigure}

@@

		\hspace*{5pt}
		% 放大的图例（占满单栏）
		\includegraphics[width=0.98\columnwidth]{figures/exp/range_legend.pdf}

		%\vspace{0.1cm}
		\begin{subfigure}[t]{0.507\columnwidth}
			\centering
			\captionsetup{font=small}
			\includegraphics[width=\linewidth]{figures/exp/range_multimodel.pdf}
			\caption{\footnotesize MultiModel}
			\label{fig:MultiModel range}
		\end{subfigure}
		\hfill
		\begin{subfigure}[t]{0.48\columnwidth}
			\centering
			\captionsetup{font=small}
			\includegraphics[width=\linewidth]{figures/exp/range_multimodel_1.pdf}
			\caption{\footnotesize MultiModel}
			\label{fig:MultiModel range}
		\end{subfigure}
		\hfill
		\begin{subfigure}[t]{0.507\columnwidth}
			\centering
			\includegraphics[width=\linewidth]{figures/exp/range_85.pdf}
			\caption{\footnotesize Different Computer}
			\label{fig:Different Computer range}
		\end{subfigure}
		\hfill
		\begin{subfigure}[t]{0.48\columnwidth}
			\centering
			\includegraphics[width=\linewidth]{figures/exp/range_deep100M.pdf}
			\caption{\footnotesize Big Dataset}
			\label{fig:Big Dataset range}
		\end{subfigure}

		\caption{\textcolor{red}TrinityExtension——Range}
		\label{fig:TrinityExtension——Range}
	\end{figure}

>>>>

		%\vspace{0.1cm}
		\begin{subfigure}[t]{0.98\columnwidth}
		\centering
		% First plot for the shared subfigure
		\includegraphics[width=0.507\linewidth]{figures/exp/attribute_multimodel.pdf}
		\hfill % Pushes the two plots apart
		% Second plot for the shared subfigure
		\includegraphics[width=0.48\linewidth]{figures/exp/attribute_multimodel_1.pdf}

		% Shared caption for the two plots above
		\captionsetup{font=small}
		\caption{MultiModel}
		\label{fig:MultiModel_range} % A single label for the combined subfigure
		\end{subfigure}

@@

	% Legend
	\includegraphics[width=0.98\columnwidth]{figures/exp/range_legend.pdf}
	\vspace{0.2cm} % Added a little space after the legend

	% --- Subfigure (a) containing two plots ---
	\begin{subfigure}[t]{0.98\columnwidth}
		\centering
		% First plot for the shared subfigure
		\includegraphics[width=0.507\linewidth]{figures/exp/range_multimodel.pdf}
		\hfill % Pushes the two plots apart
		% Second plot for the shared subfigure
		\includegraphics[width=0.48\linewidth]{figures/exp/range_multimodel_1.pdf}

		% Shared caption for the two plots above
		\captionsetup{font=small}
		\caption{MultiModel}
		\label{fig:MultiModel_range} % A single label for the combined subfigure
	\end{subfigure}

	% --- Subfigure (b) ---
	\begin{subfigure}[t]{0.507\columnwidth}
		\centering
		\includegraphics[width=\linewidth]{figures/exp/range_85.pdf}
		\captionsetup{font=small}
		\caption{Different Computer}
		\label{fig:Different_Computer_range}
	\end{subfigure}
	\hfill
	% --- Subfigure (c) ---
	\begin{subfigure}[t]{0.48\columnwidth}
		\centering
		\includegraphics[width=\linewidth]{figures/exp/range_deep100M.pdf}
		\captionsetup{font=small}
		\caption{Big Dataset}
		\label{fig:Big_Dataset_range}
	\end{subfigure}

	\caption{\textcolor{red}{TrinityExtension——Range}}
	\label{fig:TrinityExtension——Range}
\end{figure}
